Layer dependence as a measure of local dependence

### 2015-04-18 ld9.tex
More changes

The final expression in \eref{eg} can be written and approximated by, using the familiar infinitesimal notation,
\begin{equation}\label{ld}
\de \min(u,\alpha) \approx I_\alpha(u)\de \alpha  \cq I_\alpha(u) \equiv \left\{\begin{array}{cc}0\ , & u\le\alpha\ ,\\ 1\ ,& u>\alpha\ .\end{array}\right.
\end{equation}
The left hand side expression of \eref{ld} is called the $\alpha$--layer of $u$. The approximation is an equality in the limit as $\de\alpha \rightarrow 0$. The $\alpha$--layer of $u$ is an infinitesimally small increment $\de\alpha$ if $u$ exceeds $\alpha$ and is zero otherwise. Hence the $\alpha$--layer captures the movement in $u$ at $\alpha$ and ignores movements elsewhere. This interpretation is critical when forming layer dependence discussed below. \fref{flayer} illustrates the $\alpha$--layer of $u$.
\begin{figure}
  \begin{center}
    \begin{tabular}{c}
      \resizebox{100mm}{!}{\includegraphics{layer.eps}}
    \end{tabular}
    \caption{Illustration of $\alpha$-layer of $u$, written as $I_\alpha(u)\de \alpha$. The $\alpha$-layer of $u$ is only sensitive to movements in $u$ at $\alpha$ and ignores other movements.  }
    \label{flayer}
  \end{center}
\end{figure}

As often exploited  in reinsurance, any non-negative random loss or variable can be thought of as a sum of layers. Applying this concept to $u$ yields
\begin{equation}\label{decompose}
u=\int_0^u 1\de \alpha= \int_0^1 I_\alpha(u)\de\alpha    \;.
\end{equation}
Hence $u$ is formed from infinitely many layers, each capturing the movement of $u$ at a particular value.

@@

Spearman's correlation is the correlation and also the standardized covariance between two percentile rank random variables $v$ and $u$:
$$
\rho \equiv \cor(u,v) = \frac{\cov(v,u)}{\cov(u,u)}
$$
where $\cor$ and $\cov$ indicate covariance and correlation, respectively. In the final expression, the denominator $\cov(u,u)=1/12$ is a constant and a scaling factor ensuring $\rho=1$ if $u=v$ that is $u$ and $v$ are comonotonic. The numerator $\cov(v,u)$ reveals the overall dependence between $v$ and $u$, and is rewritten as, using the decomposition of $u$ in \eref{decompose},
\begin{equation}\label{spearman}
\cov(v,u)=\cov\left\{v,\int_0^1I_\alpha(u)\de \alpha \right\}= \int_0^1 \cov\{v,I_\alpha(u)\} \de \alpha \; .
\end{equation}
Therefore the covariance $\cov(v,u)$ is decomposed into infinitely many covariances $\cov\{v,I_\alpha(u)\}$ for $0\le\alpha\le 1$. Each covariance measures the dependence between $v$ and the $\alpha$--layer of $u$. In a reinsurance setting, this covariance measures dependence between a particular layer of a loss (in percentile rank terms) and another factor. Alternatively the covariance captures dependence between $v$ and movements in $u$ at $\alpha$. Both interpretations point to $\cov\{v,I_\alpha(u)\}$ as a measure of local dependence between $v$ and $u$. Similar to $\rho$, scaling this covariance with the same when $v=u$ leads to the definition of layer dependence
\begin{equation}\label{ellalpha}
\ell_\alpha \equiv \frac{\cov\{v,I_\alpha(u)\}}{\cov\{u,I_\alpha(u)\}}   \;,
\end{equation}
where the denominator 
$$
\cov\{u,I_\alpha(u)\}=\E\{uI_\alpha(u)\}-\E(u)\E\{I_\alpha(u)\}=\int_\alpha^1 u \de u - 0.5\int_\alpha^1 1 \de \alpha
$$
$$
=0.5\alpha(1-\alpha) \;.
$$
where $\E$ calculates expectations.

It is shown that if $\ell_\alpha=1$ for any $\alpha$ then $u$ and $v$ crosses $\alpha$ simultaneously: perfect dependence at $\alpha$. In addition $\ell_\alpha=1$ over $c\le\alpha\le d$ implies $u=v$ over $c\le u,v\le d$. The equality weakens as $\ell_\alpha$ decreases. These results are illustrated and formalised in \sref{sldcurve} and \sref{sdecompose} respectively.

Combining \eref{spearman} and \eref{ellalpha} yields
\begin{equation}\label{wtdavg}
\rho = \int_0^1 \ell_\alpha \frac{\cov\{u,I_\alpha(u)\}}{\cov(u,u)} \de \alpha =  \int_0^1 \ell_\alpha 6\alpha(1-\alpha) \de \alpha \; .
\end{equation}
Hence $\rho$ is a weighted average of $\ell_\alpha$ across $0\le\alpha\le 1$ with weights $w_\alpha=6\alpha(1-\alpha)$ integrating to $1$. This result is consistent with the fact that $\rho$ measures overall dependence whereas $\ell_\alpha$ decomposes $\rho$ into local dependence values. Note $w_\alpha$ has minimum $0$ at $\alpha=0$ and $1$, and increases symmetrically to maximum at $\alpha=0.5$. Hence Spearman's correlation places less emphasis on the tails which may be undesirable in finance or insurance where tail dependence is critical. Modifying the weights $w_\alpha$ leads to alternate measures of overall dependence further explored in \sref{saltoverall}.

Layer dependence summaries the dependence structure of a copula. Layer dependence provides an additional dimension of information compared to Spearman's correlation. As for any summary measure, layer dependence can mislead but is less misleading than $\rho$ and other measures of overall dependence. Layer dependence is also a more meaningful characterisation of a copula compared to the parameters of copula families such as the Clayton or Gumbel \citep{mcneil2005qrm}. Properties of and arguments for using layer dependence are explored below.

This paper assumes exchangeability: $C(u,v)=C(v,u)$ where $C$ is the copula of $(u,v)$. Hence $u$ and $v$ can be interchanged in the above expressions and those remaining in this paper. Archimedean copulas are a well known class of exchangeable copulas \citep{mcneil2005qrm}.

@@

@@

The following is an alternative expression of $\ell_\alpha$:
\begin{equation}\label{gapexp}
\ell_\alpha = \frac{\E(v|u>\alpha)-\E(v|u\leq \alpha)}{\E(u|u>\alpha)-\E(u|u\leq \alpha)}
=2 \left\{\E(v|u>\alpha)-\E(v|u\leq \alpha)\right\}\;.
\end{equation}
A proof is shown in the appendix. The middle expression in \eref{gapexp} is the expected change in $v$ relative to the expected change in $u$ when $u$ crosses $\alpha$. The latter is $0.5$ for all $\alpha$, yielding the final expression in \eref{gapexp}. Hence large $\ell_\alpha$ implies $v$ is sensitive to movements in $u$ across $\alpha$, indicating strong dependence between $v$ and $u$ at $\alpha$. When $\ell_\alpha=0$, $v$ is unchanged on average when $u$ crosses $\alpha$, hence $u$ and $v$ are not dependent at $\alpha$.

@@

The scatterplots in \fref{fillustration}  emphasize that copulas with the same overall dependence can exhibit a variety of local dependence structures and these structures are captured by layer dependence curves. Given $\alpha$, $\ell_\alpha$ is larger if scatter points are more clustered around $(\alpha,\alpha)$ and vice versa. In particular $\ell_\alpha$ increases to one in the tails of copulas exhibiting strong tail dependence. Hence $\ell_\alpha$ tracks the clustering of scatter points across the $45\circ$ line. This observation is formalised in \sref{sdecompose}.

@@

As before assume the copula of $(u,v)$ is exchangeable. The following result shows a negative relationship between layer dependence at any point and discordance and dispersion of $(u,v)$ relative to the same point. This result explains the behaviour of layer dependence curves in \fref{fillustration} where $\ell_\alpha$ increases as $(u,v)$ become scattered more closely around $(\alpha,\alpha)$, that is lower discordance and dispersion.

Layer dependence $\ell_\alpha$ can be written as
\begin{equation}\label{decomposition}
\ell_\alpha = 1-2(1+\gamma_\alpha) \delta_\alpha \;,
\end{equation}
where
$$
\gamma_\alpha \equiv \cor\{1-I_\alpha(u),I_\alpha(v)\} =\cor\{I_\alpha(u),1-I_\alpha(v)\}=\frac{\alpha^2-C(\alpha,\alpha)}{\alpha(1-\alpha)}\;,
$$
and
$$
\delta_\alpha \equiv\E\left\{(|u-v|)|(u-\alpha)(v-\alpha)<0\right\}  \; .
$$
A proof of \eref{decomposition} is shown in the appendix. The correlation $-1\leq\gamma_\alpha\leq 1$ measures the tendency for $(u,v)$ to be discordant at $\alpha$: either $u$ or $v$ is above $\alpha$ and the other is below $\alpha$. The expectation $0\leq \delta_\alpha\leq 1$ measures the average dispersion between discordant $u$ and $v$ at $\alpha$, noting $(u-\alpha)(v-\alpha)<0$ is equivalent to $u>\alpha$ and $v\le\alpha$ or $u\le\alpha$ and $v>\alpha$.

@@

The result in \eref{decomposition} explains the behaviour of layer dependence curves in \fref{fillustration}. Layer dependence $\ell_\alpha$ is larger if scatter points are more clustered around $(\alpha,\alpha)$, that is, fewer discordant points at $\alpha$, and discordant points at $\alpha$ are closer to the $45^\circ$ degree line. The former indicates smaller $\gamma_\alpha$ and the latter indicates smaller $\delta_\alpha$. Opposite observations apply for small $\ell_\alpha$.

@@

\begin{table}[h]
  \begin{center}
\begin{tabular}{|c|c|c|c| }
 \hline
 Copula & Description & LD \\
 \hline
 $uv$ & Independent $u$ and $v$ & 0 \\
 $\min(u,v)$ & Comonotonic $u$ and $v$ & 1 \\
 $\max(u+v-1,0)$ & Countermonotonic $u$ and $v$ & $-1$\\
 $u-C(u,1-v)$ & Replace $v$ with $1-v$ & $-\ell_\alpha$ \\
 $v-C(1-u,v)$ & Replace $u$ with $1-u$ & $-\ell_{1-\alpha}$ \\
 $u+v-1+C(1-u,1-v)$ & Replace $v$ and $u$ with complements & $\ell_{1-\alpha}$\\
 \hline
\end{tabular}
\caption{Layer dependence (LD) for various specific copulas and transformations.}\label{table1}
  \end{center}
\end{table}

@@

\section{Connections to existing measures of tail dependence}\label{sliterature}

@@

\subsection{Tail dependence coefficients}

\cite{joe1997multivariate} defines coefficients of lower and upper tail dependence where coefficients of one indicate perfect tail dependence. The following shows that layer dependence characterises perfect tail dependence equivalently as \cite{joe1997multivariate}.

Coefficients of lower and upper positive tail dependence are defined as the limiting conditional tail probabilities
$$
\lambda_L \equiv \lim_{\alpha\rightarrow 0^+} \p(v\leq \alpha |u\leq \alpha) \cq
\lambda_U \equiv \lim_{\alpha\rightarrow 1^-} \p(v>\alpha |u>\alpha) \;.
$$
Unit coefficients indicate perfect positive tail dependence, and occur if and only if $u$ and $v$ converge simultaneously to 0 (lower tail) or 1 (upper tail). Coefficients of negative tail dependence replace $v\leq \alpha$ and $v>\alpha$ in the above expressions with $v>1-\alpha$ and $v\leq 1-\alpha$, respectively.  \cite{sweeting2013calculating} discusses the drawback of these coefficients and suggests a modification by weakening the limits, yielding links to tail concentration functions discussed below.

The following shows $\lambda_L=1$ is equivalent to $\ell_0=1$ and $\lambda_U=1$ is equivalent to $\ell_1=1$. Similar links apply to negative tail dependence. Hence \cite{joe1997multivariate} and layer dependence characterise perfect tail dependence equivalently. From \eref{gapexp}, $\ell_1=2\E(v|u=1)-1$ implying $\ell_1=1$ if and only if $\E(v|u=1)=1$. Hence $\ell_1=1$ if and only if $u=1$ implies $v=1$, which is equivalent to $\lambda_U=1$. In addition $\ell_0=1$ if and only if $u=0$ implies $v=0$, which is equivalent to $\lambda_L=1$. Similar proofs apply to perfect negative tail dependence.

@@

Tail concentration \citep{venter2002tails} is a local dependence measure formed from lower and upper conditional tail probabilities. Similar to layer dependence curves, tail concentration functions characterise the dependence structure of copulas and are used to identify families of copulas. The following shows an intimate connection between layer dependence and tail concentration. In addition layer dependence refines tail concentration by incorporating additional information on dispersion defined in \sref{sdecompose}.

The tail concentration function at $\alpha$ is:
$$
\tau_\alpha  \equiv \left\{\begin{array}{cc}\p(v\leq \alpha|u\leq \alpha)\ ,  & \alpha\leq 0.5\ \\
\p(v>\alpha|u>\alpha)\ ,& \alpha>0.5 \ \end{array}\right.
=\left\{\begin{array}{cc}\frac{C(\alpha,\alpha)}{\alpha}\ , & \alpha\leq 0.5\ \\
\frac{1-2\alpha+C(\alpha,\alpha)}{1-\alpha}\ ,& \alpha>0.5 \ \end{array}\right. \;.
$$
Higher $\tau_\alpha$ implies $u$ and $v$ are more likely to fall in the same tail -- lower tail if $\alpha\leq 0.5$ and upper tail if $\alpha>0.5$. In addition taking limits $\alpha\rightarrow 0$ and $\alpha\rightarrow 1$ of $\tau_\alpha$ yield coefficients of tail dependence $\lambda_L$ and $\lambda_U$ discussed above. Properties of tail concentration and its applications to distinguish families of copulas are further discussed in \cite{durante2014copulas}.

To show the connection between layer dependence and tail concentration, first standardise $\tau_\alpha$ by subtracting its value under independence and dividing by the difference between $\tau_\alpha$ under comonotonicity and independence:
$$
\tau_\alpha^*=\frac{\tau_\alpha-\tau_\alpha^0}{\tau_\alpha^+-\tau_\alpha^0} = \frac{C(\alpha,\alpha)-\alpha^2}{\alpha(1-\alpha)}
$$
where $\tau_\alpha^0=\tau_\alpha$ if $u$ and $v$ are independent and $\tau_\alpha^+=\tau_\alpha$ if $u=v$, hence
$$
\tau_\alpha^0=\left\{\begin{array}{cc}\alpha \ , & \alpha\leq 0.5\ \\
1-\alpha \ ,& \alpha>0.5 \ \end{array}\right.
\cq
\tau_\alpha^+=1 \;.
$$
Standardised tail concentration $\tau_\alpha^*$ is 1 if $u=v$ and $0$ if $u$ and $v$ are independent.

Note the final expression for $\tau_\alpha^*=-\gamma_\alpha$ where $\gamma_\alpha$ measures discordance between $u$ and $v$ at $\alpha$ as discussed in \sref{sdecompose}. Combining this result with \eref{decompose} yields
$$
\ell_\alpha=1-2\delta_\alpha(1-\tau_\alpha^*) = 1-2\delta_\alpha + 2\delta_\alpha \tau_\alpha^*
$$
where $\delta_\alpha$ as defined in \sref{sdecompose} measures average dispersion between discordant $u$ and $v$ at $\alpha$. Hence given $\alpha$, $\ell_\alpha$ is increasing in $\tau_\alpha^*$ and $\tau_\alpha$. It is also straightforward to show that $\ell_\alpha=1$ is equivalent to $\tau_\alpha^*=1$.

@@

Thus $\ell_\alpha$ integrates copulas to reduce their dimension from two and one, and scales the result to ensure it lies between $\pm 1$. This computation extracts the dependence structure from the copula. In comparison, tail concentration performs the dependence extraction by computing the diagonal section $C(\alpha,\alpha)$.

@@

A direct consequence of the copula integration result \eref{copula} is layer dependence preserves convex combinations of random variables, as follows.

@@

\subsection{Layer dependence does not uniquely characterise a copula}

Layer dependence curves extract and summarise dependence information from a copula. Therefore layer dependence curves do not uniquely specify the copula, as shown in the following counterexample.

@@

Non-uniqueness is seen from another perspective using \eref{gapexp}: $\ell_\alpha$ only captures first-order conditional tail expectations. Hence two copulas with equal first-order conditional tail expectations have equal layer dependence.

\subsection{Non-exchangeability}

layer dependence retains coherence and dependence-measuring properties when the joint distribution of $(u,v)$ is non-exchangeable. However the calculation of layer dependence differs according to the conditioning variable:
$$
2\left\{\E(v|u>\alpha)-\E(v|u\leq \alpha)\right\} \neq
2\left\{\E(u|v>\alpha)-\E(u|v\leq \alpha)\right\} \;.
$$
The left expression measures local dependence with respect to $u$. The right expression measures local dependence with respect to $v$, and is not necessarily equal to the left. A similar concept applies in regression, where the regression coefficient depends on the assignment of explanatory and dependent variables.

>>>>

The final expression in \eref{eg} can be written and approximated by, using the familiar infinitesimal notation,
\begin{equation}\label{ld}
\de \min(u,\alpha) \approx I_\alpha(u)\de \alpha  \cq I_\alpha(u) \equiv \left\{\begin{array}{cc}0\ , & u\le\alpha\ ,\\ 1\ ,& u>\alpha\ .\end{array}\right.
\end{equation}
The left hand side expression in \eref{ld} is called the $\alpha$--layer of $u$. The approximation becomes exact  as $\de\alpha \rightarrow 0$. The $\alpha$--layer of $u$ is an infinitesimally small increment $\de\alpha$ if $u$ exceeds $\alpha$ and is zero otherwise. Hence the $\alpha$--layer captures the movement in $u$ at $\alpha$ and ignores movements elsewhere.  \fref{flayer} illustrates the $\alpha$--layer of $u$.
\begin{figure}
  \begin{center}
    \begin{tabular}{c}
      \resizebox{100mm}{!}{\includegraphics{layer.eps}}
    \end{tabular}
    \caption{Illustration of $\alpha$-layer of $u$, written as $I_\alpha(u)\de \alpha$. The $\alpha$-layer of $u$ is only sensitive to movements in $u$ at $\alpha$ and ignores other movements.  }
    \label{flayer}
  \end{center}
\end{figure}

As often exploited  in reinsurance, any non-negative random loss or variable can be thought of as a sum of layers:
\begin{equation}\label{decompose}
u=\int_0^u 1\de \alpha= \int_0^1 I_\alpha(u)\de\alpha    \;.
\end{equation}
Hence $u$ is formed from infinitely many $\alpha$--layers, each capturing the movement of $u$ at a different $\alpha$.

@@

Spearman's correlation is the correlation and also the standardized covariance between two percentile rank random variables $v$ and $u$:
\begin{equation}\label{rho}
\rho \equiv \cor(u,v) = \frac{\cov(v,u)}{\sqrt{\cov(v,v)\cov(u,u)}}=\frac{\cov(v,u)}{\cov(u,u)} = 12\cov(v,u)\ ,
\end{equation}
where $\cor$ and $\cov$ indicate covariance and correlation, respectively. In this paper assume the copula of $u$ and $v$ is exchangeable \citep{nelson1999ic}.

Using the decomposition of $u$ in \eref{decompose},
\begin{equation}\label{spearman}
\cov(v,u)=\cov\left\{v,\int_0^1I_\alpha(u)\de \alpha \right\}= \int_0^1 \cov\{v,I_\alpha(u)\} \de \alpha \; .
\end{equation}
Therefore the covariance $\cov(v,u)$ is decomposed into infinitely many covariances $\cov\{v,I_\alpha(u)\}$ for $0\le\alpha\le 1$. Each covariance measures the dependence between $v$ and the $\alpha$--layer of $u$. In a reinsurance setting, this covariance measures dependence between a particular layer of a loss (in percentile rank terms) and another factor. Alternatively the covariance captures dependence between $v$ and movements in $u$ at $\alpha$. Both interpretations point to $\cov\{v,I_\alpha(u)\}$ as a measure of local dependence between $v$ and $u$. Similar to $\rho$, scaling this covariance with the same when $v=u$ leads to the definition of layer dependence
\begin{equation}\label{ellalpha}
\ell_\alpha \equiv \frac{\cov\{v,I_\alpha(u)\}}{\cov\{u,I_\alpha(u)\}}   \;,
\end{equation}
where the denominator
\begin{equation}\label{covui}
\cov\{u,I_\alpha(u)\}=\E[\{u-\E(u)\}I_\alpha(u)]=\int_\alpha^1 \left(u  - \frac{1}{2}\right)\de \alpha = \frac{\alpha(1-\alpha)}{2}\ .
\end{equation}
Combining \eref{rho}, \eref{spearman}, \eref{ellalpha}  and \eref{covui} yields
\begin{equation}\label{wtdavg}
\rho = 12\int_0^1 \ell_\alpha \cov\{u,I_\alpha(u)\} \de \alpha =  \int_0^1 \ell_\alpha 6\alpha(1-\alpha) \de \alpha \; .
\end{equation}

Hence $\rho$ is a weighted average of $\ell_\alpha$ for $0\le\alpha\le 1$ with weights $w_\alpha=6\alpha(1-\alpha)$ integrating to $1$, and $\ell_\alpha$ decomposes $\rho$ into local dependence values. Note $w_\alpha$ has minimum $0$ at $\alpha=0$ and $1$ and increases symmetrically to maximum at $\alpha=0.5$. Hence Spearman's correlation places little emphasis on the tails which may be undesirable in finance or insurance where tail dependence is critical. Modifying the weights $w_\alpha$ leads to alternate measures of overall dependence further discussed in \sref{saltoverall}.

Layer dependence summarises the dependence structure of a copula. Layer dependence provides additional information compared to Spearman's correlation: how dependence varies across the joint distribution. As  any summary measure, layer dependence can mislead but is less misleading than $\rho$ and other measures of overall dependence. Layer dependence is  a more meaningful characterisation of a copula compared to the parameters of copula families such as the Clayton or Gumbel \citep{mcneil2005qrm}. Properties of and arguments for using layer dependence are explored below.

\begin{comment}
(** Not sure this is needed **) This paper assumes exchangeability: $C(u,v)=C(v,u)$ where $C$ is the copula of $(u,v)$. Hence $u$ and $v$ can be interchanged in the above expressions and those remaining in this paper. Archimedean copulas are a well known class of exchangeable copulas \citep{mcneil2005qrm}.
\end{comment}

@@

\begin{comment}
Similar to correlation, $0\le\ell_\alpha\le 1$ with $\ell_\alpha=-1$, $0$ and $1$ if $u$ and $v$ are countermonotonic, independent and comonotonic, respectively. In addition $\ell_\alpha$ increases with the correlation order of $(u,v)$. Replacing $u$ or $v$ with their complement leads to straightforward changes in layer dependence.

Layer dependence $\ell_\alpha$ satisfies  coherence properties for all $0\le\alpha\le 1$. These properties are similar to those of Spearman's correlation and are formalised in \sref{scoherence}.
\end{comment}

@@

In \sref{sldcurve} and \sref{sdecompose} it is illustrated and shown  that  $\ell_\alpha=1$ for a particular $\alpha$ occurs if and only if $I_\alpha(u)$ and $I_\alpha(v)$ are both simultaneously 1 or 0: that $u$ and $v$ are both  either greater than $\alpha$ or less than $\alpha$.  Additionally $\ell_\alpha=1$ for $c\le\alpha\le d$ implies $u=v$ over $c\le u,v\le d$. The results weaken as $\ell_\alpha$ decreases from 1.

The following is an alternative expression for $\ell_\alpha$:
\begin{equation}\label{gapexp}
\ell_\alpha = \frac{\E(v|u>\alpha)-\E(v|u\leq \alpha)}{\E(u|u>\alpha)-\E(u|u\leq \alpha)}
=2 \left\{\E(v|u>\alpha)-\E(v|u\leq \alpha)\right\}\;.
\end{equation}
Proofs are in an appendix. The middle expression in \eref{gapexp} is the expected change in $v$ relative to the expected change in $u$ when $u$ crosses $\alpha$. The latter is $0.5$ for all $\alpha$, yielding the final expression in \eref{gapexp}. Hence large $\ell_\alpha$ implies $v$ is sensitive to movements in $u$ across $\alpha$, indicating strong dependence between $v$ and $u$ at $\alpha$. When $\ell_\alpha=0$, $v$ is unchanged on average when $u$ crosses $\alpha$, hence $u$ and $v$ are not dependent at $\alpha$.

@@

The scatterplots in \fref{fillustration}  emphasize that copulas with the same overall dependence can exhibit a variety of local dependence structures and these structures are captured by layer dependence curves. Given $\alpha$, $\ell_\alpha$ is larger if scatter points are more clustered around $(\alpha,\alpha)$ and vice versa. In particular $\ell_\alpha$ increases to one in the tails of copulas exhibiting strong tail dependence. Hence $\ell_\alpha$ tracks the clustering of scatter points across the $45^\circ$ line. This  is formalised in \sref{sdecompose}.

@@

Again assuming the copula $C$ of $(u,v)$ is exchangeable, then layer dependence $\ell_\alpha$ can be written as
\begin{equation}\label{decomposition}
\ell_\alpha = 1-2(1+\gamma_\alpha) \delta_\alpha \;,
\end{equation}
where
$$
\gamma_\alpha \equiv \cor\{1-I_\alpha(u),I_\alpha(v)\} =\cor\{I_\alpha(u),1-I_\alpha(v)\}=\frac{\alpha^2-C(\alpha,\alpha)}{\alpha(1-\alpha)}\;,
$$
and (** can the next conditioning be written in terms of $I_\alpha(u)$ etc the only way I can think of is $|I_\alpha(u)-I_\alpha(v)|=1$ which is equally messy?**)
$$
\delta_\alpha \equiv\E\left\{(|u-v|)|(u-\alpha)(v-\alpha)<0\right\}  \; .
$$
A proof of \eref{decomposition} is given in the appendix. Equation \eref{decomposition}, as shown below, explains the behaviour of layer dependence curves in \fref{fillustration} where $\ell_\alpha$ for any $\alpha$ increases as the scatter become more tightly concentrated  around the point $(\alpha,\alpha)$, that is lower discordance and dispersion relative to the same point.

The correlation $-1\leq\gamma_\alpha\leq 1$ measures the tendency for $(u,v)$ to be discordant at $\alpha$: either $u$ or $v$ is above $\alpha$ and the other is below $\alpha$. The expectation $0\leq \delta_\alpha\leq 1$ measures the average dispersion between discordant $u$ and $v$ at $\alpha$, noting $(u-\alpha)(v-\alpha)<0$ is equivalent to $u>\alpha$ and $v\le\alpha$ or $u\le\alpha$ and $v>\alpha$.

@@

Applying \eref{decomposition} and generalising the illustration in \fref{finterpretation} explains the behaviour of layer dependence curves in \fref{fillustration}. Layer dependence $\ell_\alpha$ is larger if scatter points are more clustered around $(\alpha,\alpha)$, that is, fewer discordant points at $\alpha$, and discordant points at $\alpha$ are closer to the $45^\circ$ degree line. The former indicates smaller $\gamma_\alpha$ and the latter indicates smaller $\delta_\alpha$. Opposite observations apply for small $\ell_\alpha$.

@@

\begin{table}[h]
  \begin{center}
\begin{tabular}{c|c|c }
 Copula & Description & LD \\
 \hline
 $uv$ & Independent $u$ and $v$ & 0 \\
 $\min(u,v)$ & Comonotonic $u$ and $v$ & 1 \\
 $\max(u+v-1,0)$ & Countermonotonic $u$ and $v$ & $-1$\\
 $u-C(u,1-v)$ & Replace $v$ with $1-v$ & $-\ell_\alpha$ \\
 $v-C(1-u,v)$ & Replace $u$ with $1-u$ & $-\ell_{1-\alpha}$ \\
 $u+v-1+C(1-u,1-v)$ & Replace $v$ and $u$ with complements & $\ell_{1-\alpha}$\\
 \hline
\end{tabular}
\caption{Layer dependence (LD) for different copulas and transformations.}\label{table1}
  \end{center}
\end{table}

@@

\section{Connections to measures of tail dependence}\label{sliterature}

@@

\subsection{Tail dependence}

\cite{joe1997multivariate} defines coefficients of lower and upper tail dependence as
$$
\lambda_L \equiv \lim_{\alpha\downarrow 0} \p(v\leq \alpha |u\leq \alpha) \cq
\lambda_U \equiv \lim_{\alpha\uparrow 1} \p(v>\alpha |u>\alpha) \;.
$$
Unit coefficients indicate perfect positive tail dependence, and occur if and only if $u$ and $v$ converge simultaneously to 0 (lower tail) or 1 (upper tail). Coefficients of negative tail dependence replace $v\leq \alpha$ and $v>\alpha$ in the above expressions with $v>1-\alpha$ and $v\leq 1-\alpha$, respectively.  \cite{sweeting2013calculating} discusses the drawback of these coefficients and suggests a modification by weakening the limits, yielding links to tail concentration functions discussed below.

(** I'm not completely comfortable with this **) The following shows $\lambda_L=1$ is equivalent to $\ell_0=1$ and $\lambda_U=1$ is equivalent to $\ell_1=1$.   This follows since  $\ell_1=1$ if and only if $\E(v|u=1)=1$. Hence $\ell_1=1$ if and only if $u=1$ implies $v=1$, which is equivalent to $\lambda_U=1$.
From \eref{gapexp}, $\ell_1=2\E(v|u=1)-1$ implying In addition $\ell_0=1$ if and only if $u=0$ implies $v=0$, which is equivalent to $\lambda_L=1$. Similar proofs apply to perfect negative tail dependence.
Hence the measures of  \cite{joe1997multivariate} and layer dependence characterise perfect tail dependence equivalently.

@@

Tail concentration \citep{venter2002tails} is a local dependence measure formed from lower and upper conditional tail probabilities. Similar to layer dependence curves, tail concentration functions describe the dependence structure of copulas and are used to identify families of copulas. The following shows an intimate connection between layer dependence and tail concentration. In addition layer dependence refines tail concentration by incorporating additional information on dispersion defined in \sref{sdecompose}.

The tail concentration function at $\alpha$ is:
$$
\tau_\alpha  \equiv \left\{\begin{array}{cc}\p(v\leq \alpha|u\leq \alpha)\ ,  & \alpha\leq 0.5\ \\
\p(v>\alpha|u>\alpha)\ ,& \alpha>0.5 \ \end{array}\right.
=\left\{\begin{array}{cc}\frac{C(\alpha,\alpha)}{\alpha}\ , & \alpha\leq 0.5\ \\
\frac{1-2\alpha+C(\alpha,\alpha)}{1-\alpha}\ ,& \alpha>0.5 \ \end{array}\right. \;.
$$
Higher $\tau_\alpha$ implies $u$ and $v$ are more likely to fall in the same tail -- lower tail if $\alpha\leq 0.5$ and upper tail if $\alpha>0.5$. In addition taking limits $\alpha\rightarrow 0$ and $\alpha\rightarrow 1$ of $\tau_\alpha$ yield coefficients of tail dependence $\lambda_L$ and $\lambda_U$ discussed above. Properties of tail concentration and its applications to distinguish families of copulas are  discussed in \cite{durante2014copulas}.

To show the connection between layer dependence and tail concentration, first standardise $\tau_\alpha$ by subtracting $\tau_\alpha^0$, its value under independence and dividing by the difference between $\tau_\alpha^+$, its value if $u=v$, and $\tau_\alpha^0$:
$$
\tau_\alpha^*=\frac{\tau_\alpha-\tau_\alpha^0}{\tau_\alpha^+-\tau_\alpha^0} = \frac{C(\alpha,\alpha)-\alpha^2}{\alpha(1-\alpha)}
$$
where
$$
\tau_\alpha^0=\left\{\begin{array}{cc}\alpha \ , & \alpha\leq 0.5\ \\
1-\alpha \ ,& \alpha>0.5 \ \end{array}\right.
\cq
\tau_\alpha^+=1 \;.
$$
Then $\tau_\alpha^*$ equals 1 or 0, if $u=v$ or  if $u$ and $v$ are independent, respectively.

Note the final expression for $\tau_\alpha^*=-\gamma_\alpha$ where $\gamma_\alpha$ measures discordance between $u$ and $v$ at $\alpha$ as discussed in \sref{sdecompose}. Combining this result with \eref{decomposition} yields
$$
\ell_\alpha=1-2\delta_\alpha(1-\tau_\alpha^*) = 1-2\delta_\alpha + 2\delta_\alpha \tau_\alpha^*
$$
where $\delta_\alpha$ as defined in \sref{sdecompose} measures average dispersion between discordant $u$ and $v$ at $\alpha$. Hence given $\alpha$, $\ell_\alpha$ is increasing in $\tau_\alpha^*$ and $\tau_\alpha$. It is also straightforward to show that $\ell_\alpha=1$ is equivalent to $\tau_\alpha^*=1$.

@@

Thus $\ell_\alpha$ summarises a copula: reducing the dimension from two and one, and scales the result to ensure it lies between $\pm 1$. This computation extracts the dependence structure from the copula. In comparison, tail concentration performs the dependence extraction by computing the diagonal section $C(\alpha,\alpha)$.

@@

A direct consequence of the copula integration result \eref{copula} is layer dependence preserves convex combinations of copulas, as follows.

@@

\begin{comment}
\subsection{Layer dependence does not uniquely characterise a copula}

(** Not sure this section is needed **)  Layer dependence curves extract and summarise dependence information from a copula. Therefore layer dependence curves do not uniquely specify the copula, as shown in the following counterexample.

@@

Non-uniqueness is seen from another perspective using \eref{gapexp}: $\ell_\alpha$ only captures first-order conditional tail expectations. Hence two copulas with equal first-order conditional tail expectations have equal layer dependence.
\end{comment}
